% Manually putting in the title here and using donotrepeattitle in paper.tex so that I don't have a section header for "Introduction", but there is still a section in the VScode navigation for all the intro subsections to live within. 
\makeatletter
\section{\@title}
\makeatother

Legislative redistricting is the process of drawing district boundaries for elections. In the United States, the redrawing of congressional district boundaries takes place every ten years, after the decennial census. This process of redistricting is generally performed by state legislatures, hybrid commissions, or independent commissions \parencite{princetongerrymanderingproject}. 

When redistricting congressional districts, the two most fundamental requirements are that the districts must be "reasonably compact" and of equal population, meaning less than 1\% deviation from the average district population \parencite{1964}. 

Gerrymandering is drawing the electoral district boundaries in such a way as to unfairly benefit a particular interest group. The term originated from the "salamander-shaped" districts drawn for the Massachusetts state senate in 1812 \parencite{hunter2011}. There are two primary methods of gerrymandering: "packing" and "cracking." The US Supreme Court defines these as creating districts where the opposing party holds far above the majority or is spread out amongst so many districts as to become ineffectual, respectively \parencite{1986}.

Automated redistricting algorithms are one proposed method to prevent gerrymandering. As early as 1961, such "automatic and impersonal procedure" were thought to be able to "produce results not markedly inferior to those which would be arrived at by a genuinely disinterested commission" \parencite[110]{vickrey1961}.

Broadly speaking, the goal of such algorithms is to create a district map (or a selection of maps) that is deemed "fair." Some algorithms obey only equal population and compactness constraints \parencite[e.g.][]{hu1995, altman2009, chen2013}, while more-modern algorithms have evolved to address real-world considerations such as the stipulations of the Voting Rights Act, incumbency, and the competitiveness of districts \parencite[e.g][]{lara-caballero2019, fifield2020, mccartan2020}. 

In parallel to the development of these algorithms, measures of compactness and of partisan fairness have also been developed. These compactness measures include measures evaluating the geometric compactness of districts \parencite[e.g.][]{polsby1991, schwartzberg1966,harris1964, maceachren1985, reock1961, boyce1964}, as well as population-weighted compactness \parencite[e.g][]{fryer2007} and compactness from a graph-theory perspective \parencite[e.g.][]{dube2016}. Proposed measures of partisan fairness generally try to characterize the relationship between the average proportion of votes won by a party in each district and the proportion of seats acquired in the delegation or legislature. Some measures seek to simulate the power balance in the legislature at varying average district votes \parencite[e.g.][]{tufte1973}, while others aim to compute a single number to describe the fairness of a redistricting plan provided the results of a single election \parencite[e.g.][]{stephanopoulos2014,katz2020,warrington2018,mcdonald2015,wang2016}. Neither the scholarly literature nor the courts agree on a single correct compactness or partisan fairness measure. 

\subsection{Overview of Method}

While the scholarship is focused on the development of new automated redistricting algorithms \parencite{fifield2020},  the evaluation of the mathematical rigor of proposed metrics \parencite[see][]{katz2020}, and on the detection of gerrymandering in existing redistricting plans \parencite[e.g.][]{herschlag2017, duchin2018a}, research into the empirical performance of these algorithms in a real-world redistricting scenario is lacking \parencite{fifield2020a}.

Identifying this gap leads me to my research question, \emph{How do the hypothetical district maps for the Virginia Congressional delegation for the 2020s generated by different automated redistricting algorithms compare based on partisan fairness and compactness measures?}

Now I will justify the components of my question. I chose Virginia as my case state because it 
\begin{seriate}
    \item has had to redistrict mid-decade due to gerrymandering \parencite[see][]{2016a},
    \item is a competitive state from an electoral perspective, 
    \item recently approved a constitutional amendment that institutes a redistricting process driven by a hybrid commission \parencite{ballotpediastaff2020}, and
    \item has a complex geography.
\end{seriate}
This complex history with redistricting and the shift towards an external commission means that varying sets of maps will be proposed, and some will likely have been generated by automated algorithms.\footnote{I choose to simulate the redistricting of the 11 congressional districts as opposed to the 100 House of Delegates districts as the latter was computationally prohibitive.}

With the goal of comparing the performance of different automated redistricting algorithms, I chose two algorithms of different levels of complexity that could both meet the real-world requirements of equal population and compactness. The Compact Random Seed Growth algorithm (CRSG) \parencite{chen2013} and the Sequential Monte Carlo algorithm (SMC) \parencite{mccartan2020} were chosen as they satisfy the aforementioned requirement, are implemented by an open-source software package \parencite{fifield2020d}, and represent the evolution in the scholarship on such algorithms. 

Since there is no single measure of partisan fairness nor compactness agreed upon by the scholarship, I selected representative samples of the different perspectives on each measure. For compactness, one geometry-based measure (the Polsby-Popper score \parencite{polsby1991}) and one graph-theory-based measure (the edge-cut compactness score \parencite{dube2016}) were chosen. For partisan fairness, the simulated election measure (the seats-votes curve \parencite{tufte1973}) and four single-valued fairness measures (partisan bias \parencite{katz2020}, the efficiency gap \parencite{stephanopoulos2014}, declination \parencite{warrington2018}, and the mean-median difference \parencite{mcdonald2015}) were chosen. Together, these five partisan fairness measures represent differing and disputed views on how "fairness" should be quantified \parencite{katz2020}. 

I have chosen to conduct a case-study using the experimental research method, as it allows me to infer the causality of the different redistricting algorithms on the redistricting plan outcomes. 

This paper begins with a \hyperref[sec:litreview]{literature review} of the scholarship on automated redistricting algorithms, compactness measures, and partisan fairness measures. Further details of my chosen research method are explained in the \hyperref[sec:method]{Method} section. The results of my analysis are presented, interpreted, and reflected upon in the \hyperref[sec:results]{Results}, \hyperref[sec:disc]{Discussion}, and \hyperref[sec:conc]{Conclusion} sections, respectively. 